TESSELLATE asteroids paper

### 2025-09-16 PaperDraft/paperDraft.tex
figures added

From their lightcurves, the rotation period of the object can be determined by using a Lomb-Scargle periodogram \citep[\citet{Lomb1976, Scargle1982}, but see][ for a review]{VanderPlas2018}.
The new \ttt{NIFTY-LS} package \citep{Garrison2024},  as implemented in \ttt{astropy} \citep{Astropy2013,Astropy2018,Astropy2022}, was used to calculate the periodogram.
The periodogram searched for rotation periods between the Nyquist limit of \qty{20}{\minute} and a maximum of \qty{17}{\day} \citep[the value used in][ due to the length of the lightcurve]{McNeill2023}.

@@

An example of the periodogram fit to the lightcurve is given by \autoref{fig:perEx}. 
A clear maximum power is found in the top panel, and the associated frequency is fed to the model shown in the middle panel.
The residuals in the lower panel are small for the most part, with large deviations from 0 only where the lightcurve has large uncertainty in its flux. 
Thanks to the known frequency from LCDB plotted, we note that we recover the double frequency alias, this is explored further below.

@@

Not every lightcurve and periodogram is good enough for analysis, so some quality cuts are applied.
A minimum brightness cut of 10 counts $\sim \qty{18}{mag}$ %?which way around is better? 
is applied, otherwise the lightcurve is indistinguishable from noise. %* This is emperical from match to LCDB, should that fig be here?
The periodograms could return the \qty{17}{\day} maximum value if no physical period was detected, so following \citet{McNeill2023} we discard any periods within \qty{10}{\percent} of this maximum. %TODO min period of \qty{1}{\hour} 
%! This and the doubling of the periods means we can't get fast rotators at all. Need to think this through
Also following \citeauthor{McNeill2023}, we do not trust the periods of any lightcurve with $\leq 200$ datapoints.
%TODO are there more checks

@@

%TODO comp to LCDB
\red{57426 total, 1530 in LCDB, 868 of those pass check, 581 within 5 percent of y=2x, 28 within 5 percent of y=x}

@@

@@

We do not find any fast rotators in our dataset.
While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discovered above the barrier are small%TODO CITE Warner2009 LCDB?
, and therefore dim.
The shallow magnitude of \tessellate\ \citep{TESSELLATE} makes it difficult for dim planetesimals to pass the quality checks.

>>>>

From their lightcurves, the rotation period of the object can be determined by using a Lomb-Scargle periodogram \citep[\citet{Lomb1976, Scargle1982}, but see][ for a review]{VanderPlas2018}.
The new \ttt{NIFTY-LS} package \citep{Garrison2024},  as implemented in \ttt{astropy} \citep{Astropy2013,Astropy2018,Astropy2022}, was used to calculate the periodogram.
The periodogram searched for rotation periods between the Nyquist limit of \qty{20}{\minute} and a maximum of \qty{17}{\day} \citep[the value used in][ due to the length of the lightcurve]{McNeill2023}.
The uncertainty in the frequency is calculated as half full width half maximum of the peak found in the periodogram, crudely approximating the peaks as Gaussian.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=\textwidth]{PeriodogramUlyssesPaperFigDraft.pdf}
    \caption{An example of the periodogram analysis we conduct for each object. 
    This figure uses the lightcurve of the asteroid (5254) Ulysses from sector 29. 
    Top panel: A periodogram of the lightcurve, with the known frequency and aliases indicated (from the LCDB).
    Middle panel: The lightcurve (orange error bars) and model(dashed black). 
    Lower panel: Residuals of the difference between the model and the lightcurve.}
    \label{fig:perEx}
\end{figure}

An example of the periodogram fit to the lightcurve is given by \autoref{fig:perEx}. 
A clear maximum power is found in the top panel, and the associated frequency is fed to the model shown in the middle panel.
The residuals in the lower panel are small for the most part, with large deviations from zero only where the lightcurve has large uncertainty in its flux. 
Thanks to the known frequency from LCDB plotted, we note that we recover the double frequency alias, this is explored further below.

@@

Not every lightcurve is good enough for analysis, so some quality cuts are applied to the lightcurves.
A minimum brightness cut of 10 counts ($m_T \sim \qty{18}{mag}$) %?which way around is better? 
is applied, otherwise the lightcurve is indistinguishable from noise. %* This is emperical from match to LCDB, should that fig be here?
Following \citeauthor{McNeill2023}, we do not trust the periods of any lightcurve with $\leq 200$ data points.

There were also quality cuts applied to the periodograms. 
The methods would return the \qty{17}{\day} maximum value if no physical period was detected, so also following \citet{McNeill2023} we discard any periods within \qty{10}{\percent} of this maximum. %TODO min period of \qty{1}{\hour} 
They find that periods of $\leq \qty{3}{\hour}$ is also unreliable.
We are using data with a factor of three higher cadence, so we reduce this condition to $\leq \qty{1}{\hour}$ for the same number of observations sampled per rotation. 
%! This and the doubling of the periods means we can't get fast rotators at all. Need to think this through
%TODO are there more checks

@@

%TODO comp to LCDB
\red{57426 total, 1530 in LCDB, 868 of those pass check, 581 within 5 percent of y=2x, 28 within 5 percent of y=x}
%TODO say we think only 70% are accurate.

We present double the periods we detect from periodograms as the rotation rate of these objects.
This is because $\approx \qty{70}{\percent}$ of the asteroid that pass the quality checks in \autoref{fig:LCDBComp} line within \qty{5}{\percent} of the $y=2x$ line; they have double peaked lightcurves.
A few percent have the same period detected as in the LCDB, but a larger fraction have periods detected with no relation to their known period.
Extrapolating this trend to the \red{4328} asteroids that pass the checks, we present an accurate rotation period for $\sim \red{3030}$ of them (unfortunately we cannot say which ones).

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{LCDBCompTotalPaperFigDraft.pdf}
    \caption{Comparison of the detected period and known period for all the asteroids analysed that are in the LCDB.
    Those that failed the quality check are shown with crosses, and those that passed are filled circles. They are coloured by the measured mean \tess\ magnitude ($m_T$).
    The 1:1 correspondence line (blue), and the factor of two aliases (purple ($\times 2$) and red ($\times 0.5$)) are plotted as dotted lines.
    }
    \label{fig:LCDBComp}
\end{figure}

@@

\autoref{fig:otherWorksComp}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{M23andP20CompPaperFigDraft.pdf}
    \caption{Comparing the rotation period of asteroids found in this work to asteroids in other studies that used TESS to investigate asteroids. 
    Left: The comparison to \citet{McNeill2023} plotted as pink crosses, there are \red{139} objects.
    Right: The comparison to \citet{Pal2020} plotted as purple pluses, there are \red{117} objects. 
    Those planetesimals found in all three works are circled (\red{54} asteroids).
    Lines of 1:1 agreement are plotted (dotted blue).}
    \label{fig:otherWorksComp}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Diam-FreqPlotThisWorkPaperFigDraft.pdf}
    \caption{A rotation rate against size plot after the LCDB \citep{Warner2009} (grey
circles) with this work's addition (blue squares). The
diameters are from NEOWISE \citep{Masiero2011,NEOWISE2019} where available. The red dashed lines represent the spin barrier of \citep{Pravec2000}.}
    \label{fig:FreqDiam}
\end{figure}

We do not find any fast rotators in our dataset.
This is encapsulated by none of our objects being above the spin barrier shown in \autoref{fig:FreqDiam}.

While our Nyquist limit is well above the spin barrier of \citet{Pravec2000} thanks to \tess's fast cadence, most asteroids discovered above the barrier are small%TODO CITE Warner2009 LCDB?
, and therefore dim.
The shallow magnitude of \tessellate\ \citep{TESSELLATE} makes it difficult for dim planetesimals to pass the quality checks.
